---
title: "data_manipulation"
author: "Bingyu Sun"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import my csv files as an example.

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
                        col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv(file = "./data/FAS_pups.csv",
                        col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select variables

By listing

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, group, litter_number, gd0_weight, pups_survive)
```

by specifying 

```{r}
select(litters_data, gd_of_birth:pups_survive)

select(litters_data, group, gd_of_birth:pups_survive)
```

by saying what you want to remove:

```{r}
select(litters_data, -litter_number)
```

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)

rename(litters_data, gest_day_0_weight = gd0_weight)
```
do select and rename at the same time

use select helpers:
```{r}
select(litters_data, starts_with("gd"))

select(litters_data, litter_number, starts_with("gd"))

select(litters_data, litter_number, everything())
```
put litter_number first, then everything

#### Excercise 
 
```{r}
pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## Filter Observation

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight > 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low8")
filter(litters_data, group == "Low8" | group == "Low7")
filter(litters_data, group %in% c("Low7", "Low8"))
```

Excercise
In the pups data:
Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2
```{r}
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```

## Mutate

Create variables

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
```
create new variable, override group to a new version, make group lower case

```{r}
mutate(litters_data,
      wt_gain = (gd18_weight - gd0_weight),
      wt_gain_squared = wt_gain^2)
```

## Arrange

Arrange the data..
```{r}
arrange(litters_data, gd0_weight)
```
gd0_weight organized from smallest to large

```{r}
arrange(litters_data, pups_born_alive, gd0_weight)

arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

## Piping

Look at intermediate object approach:

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%

```
pipe 

Illustrate place holder use with 



