---
title: "Data import"
author: "Bingyu Sun"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
load tidyverse loads a bunch of packages, include read csv etc.

## Import FAS csv files

Import my first csv.

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```
use janitor package clean_name function, do not load the entire janitor package becuase if janitor and tidyverse have same function, R is gonna give you an error. So only load package that we are gonna use extensively.


## Look at data

Look at litters data.

```{r}
head(litters_data)
tail(litters_data)

skimr::skim(litters_data)
```
skimr package skim function gives info on data
also can use View function in console, don't put View function in rmd cuz gonna give error.
Steps: Import data, clean variable names, and look at it.


Skip some rows; omit variable names

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  skip = 10, col_names = FALSE)
```


Look at pups data

```{r}
pups_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
skimr::skim(pups_data)
```
Specify column types focus column to what you specifed, becasue R gnerally looks at 1000 data to determine type, if you have diff type in later data, you need to specify data type. 

## Other formats

Read in mlb data

```{r}
mlb_data = readxl::read_excel(path = "./data/mlb11.xlsx")

mlb_subset = readxl::read_excel(path = "./data/mlb11.xlsx", 
                                range = "A1:E17")
```

Read in pulse data:

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat")
```


## Compare with base R

```{r}
pups_readr = read_csv("./data/FAS_pups.csv")
pups_baser = read.csv("./data/FAS_pups.csv")

pups_readr
pups_baser$S
```
read.csv gives more data and can guess, don't use read.csv
